
# A Steenekamp Reproducible Research
## Loading and preprocessing the data
### Read the data
```{r read}
dat1<-read.csv("activity.csv")
```
### Load libraries used in Assignment
```{r Load libraries}
library(dplyr)
library(ggplot2)
library(xtable)
library(knitr)
library(chron)
```

## What is mean total number of steps taken per day?
### Histogram
```{R total number of steps}
dat2<-dat1
# Omit NA's
dat2<-na.omit(dat2)
# Group the total steps by date and calculate the sum of each day
totsteps<-dat2 %>% group_by(date) %>% summarise(tot=sum(steps))%>%arrange(date)
# Transform the date
totsteps<-totsteps %>% mutate(date = as.Date(date))
# Transform the steps
totsteps<-totsteps %>% mutate(tot = as.numeric(tot))
#Plot the Histogram
plot(totsteps$date,totsteps$tot,main="Total steps per day without imputed NA's",xlab="Date",ylab="Total Steps",type="h",col="blue")
```


### The mean of steps per day
```{r calculate Mean}
dat2<-dat1
# Calculate the mean of steps per day
meansteps<-dat2 %>% mutate(date=as.Date(date),steps=as.numeric(steps))
meansteps<-meansteps %>% group_by(date) %>% summarise(avg=mean(steps))%>%arrange(date)
meansteps<-as.data.frame.AsIs(meansteps)
meansteps
```


### The median of steps per day (not sure if the median is zero, but running out of time...)
```{r median1}
dat2<-dat1
# Calculate the median of steps per day
mediansteps<-dat2 %>% mutate(date=as.Date(date),steps=as.numeric(steps))
mediansteps<-mediansteps %>% group_by(date) %>% summarise(med=median(steps))%>%arrange(date)
mediansteps<-as.data.frame.AsIs(mediansteps)
mediansteps

```

##What is the average daily activity pattern?
### Time-series Plot
```{r Timeseries plot}
# Create a dataset for average steps per interval
avgstepsint<-dat2
# Omit NA's
avgstepsint<-na.omit(avgstepsint)
# Group and Summarise
avgstepsint<-avgstepsint %>% group_by(interval) %>% summarise(avg=mean(steps))%>%arrange(interval)
# Plot
plot(avgstepsint$interval,avgstepsint$avg,main="Time series average steps",xlab="Interval",ylab="Avg Steps",type="l",col="steelblue4")
``` 

### Which 5-minute interval contains the max number of steps? (Answer: 835)
```{r max steps}
# Get the maximum steps
max(avgstepsint$avg)
# Determine which row number
which.max(avgstepsint$avg)
# Get the interval
maxint<-avgstepsint[104,1]
maxint
```
## Imputing missing values
### The strategy is to impute the mean of all means
```{r Strategy for NAs}
# Calculate the number of NA's
dat1na<-sum(is.na(dat1))
dat1na
# Determine that only steps has NA's, none of the other variables
dat2na<-sum(is.na(dat2$steps))
dat2na
# Calculate the average number of steps for all days
dat3<-dat1
dat3<-na.omit(dat3)
dat3<-dat3 %>% group_by(date) %>% summarise(avg=mean(steps))%>%arrange(date)
meanavg<-mean(dat3$avg)
# Impute into dat4 dataset
dat4<-dat1
dat4[is.na(dat4)] = meanavg
```
### Histogram with imputed values

```{r Histogram with imputed values}
# Group and summarise imputed dataset
totisteps<-dat4 %>% group_by(date) %>% summarise(tot=sum(steps))%>%arrange(date)
# Mutate to date
totisteps<-totisteps %>% mutate(date = as.Date(date))
# Calculate the average number of steps for all days
totisteps<-totisteps %>% mutate(tot = as.numeric(tot))
# Impute into dat4 dataset
plot(totisteps$date,totisteps$tot,main="Total steps per day with imputed NA's",xlab="Date",ylab="Total Steps",type="h",col="red")
```    

## Are there differences in activity patterns between weekdays and weekends?
```{r weekends and weekdays}
# Determine weekend or not
dat5<-dat1
dat5<-na.omit(dat5)
dat5<-mutate(dat5,iswkend=is.weekend(dat5$date))
# Filter weekends
wkends<-filter(dat5,iswkend)
# Filter weekdays
wkdays<-filter(dat5,!iswkend)
# Determine averages
meanintwkd<-wkdays %>% group_by(interval) %>% summarise(avg=mean(steps))%>%arrange(interval)
meanintwke<-wkends %>% group_by(interval) %>% summarise(avg=mean(steps))%>%arrange(interval)
# Plot the graph
par(mfrow = c(1,2))
plot(meanintwkd$interval, meanintwkd$avg, type = "l", ylab = "Steps on Weekdays", xlab = "Interval",main="Weekdays",col="red")
plot(meanintwke$interval, meanintwke$avg, type = "l", ylab = "Steps on Weekends", xlab = "Interval",main="Weekends",col="blue") 
```   
